Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-15 special-issue-ci/Response to reviewers.tex

\documentclass[preprint]{elsarticle}
\biboptions{round, numbers}
\usepackage[latin1]{inputenc}
%\usepackage[T1]{fontenc}
%\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{color}
%\usepackage{setspace}
\usepackage{url}
\usepackage[english]{babel}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Volunteers in the Clouds: an Architecture for Low Cost and
  Potentially Massive Distributed Evolutionary Computation: Response to Reviewers\' comments}

\noindent
Dear Sirs,\\

We really appreciate the opportunity you have given us for improving
our work. Following the reviewers' requests and suggestions we have
deeply changed the contents, and  thus, the quality of the paper.

In the next paragraphs you can find these comments and our responses
to every one, along with the changes done.\\

\noindent
Yours sincerely,\\
The authors.

\end{document}

### 2015-11-15 special-issue-ci/Response to reviewers.tex

We really appreciate the opportunity you have given us for improving
our work. Following the reviewers' requests and suggestions we have
deeply changed the contents, and  thus, the quality of the paper.

In the next paragraphs you can find these comments and our responses
to every one, along with the changes done.\\

\noindent
Yours sincerely,\\
The authors.

\section{Response to Reviewer \#1}

\begin{quote}
The main contribution of this paper is the proposal of a Javascript based distributed system intended to
run evolutionary algorithms according to the volunteer computing scheme. The idea is that user can
contribute to the computation by using only a browser, without needing to install any plugin nor a Java
virtual machine.
The volunteer computing model appeared some years ago and it has been used in many fields. The
model is effective to solve problems demanding high computational capabilities, and at this point is where
this paper proposal introduces some concerns. The choice of Javascript code running on browsers has
some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
motivation of these kinds of distributed models is to solve complex problems, using Javascript doesnÕt
appear a right choice. ItÕs true that is a very commonly used programming language, but in the context of
Web applications, not in high performance ones. Apart from the fact of being an interpreted language
designed for the Web, it is difficult to think on a computer scientist writing a complex math code in
Javascrip. If the authors would have used as a case study a complex optimization problem then they
could have provided numerical data to sustain their claims; however, the selected problem (which should
be formulated in the paper) seems a rather academic one.
\end{quote}

\begin{quote}
The paper organization is not clear. Some comments about this:
- The title ÒVolunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
Evolutionary ComputationÓ seems to suggest that the volunteers run in the cloud, but that is not the case:
the server run􀀀 in the cloud and the volunteers do in the browsers of common PCs and laptops.
- The paper abstract does not properly summarize the paper contents. It says that ÒThat is why in this
paper we present NodIO, a client-server architecture for distributed evolutionary algorithms !Ó, but there
are two versions of the architecture, as stated later in the conclusions: ÒIn this paper two versions of a
client-server architecture for volunteer and distributed evolutionary !Ó. That second version is neither
commented in the introduction.
\end{quote}

\begin{quote}
presented in Section III, followed by the experiments and obtained results in Section IV.Ó, but Section IV
does not only contains experimental result but also the full description of the second version of the
architecture.
Summarizing the review, the authors should provide a convincing example showing the advantages of
their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
problem.
\end{quote}

\section{Response to Reviewer \#1}

\begin{quote}
Contribution:
This paper describes a client-server architecture for distributed evolutionary algorithms named NodIO.
The implementation of the architecture is written entirely in JavaScript, in order to support persistent,
asynchronous, distributed evolutionary algorithms running in the browser.
Additionally, the paper presents the results of several experiments performed to validate the proposal and
to measure the performance improvement generated by the massive parallel computing.
I think that the contribution is quite relevant, since it provides a computation platform that is flexible,
extensible and potentially massive. Personally, I consider it one of the fresher and most promising
proposals in the context of metaheuristic optimization frameworks lately. Hence, our global
recommendation of acceptance with minor changes.
\end{quote}

\begin{quote}
However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
organization and description of the platform, and the presentation and writing style.
Regarding the experimentation, I miss:
• a section discussing the threats to validity of the experiments
• an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
• A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

\begin{quote}
Regarding the organization of the paper, some important details regarding the improved framework
design named is described in the experimentation section. I would reorganize the contents to describe the
improved in detail in section 3, and focus on the experiments, its results and the discussion in section 4.
Another issue is that some details are described twice, for instance, the advantages of the use of web
workers. Fixing this issue can reduce the paper size and improve its readability.
Regarding the presentation:
• Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
• The axis labels and legends in most of the figures are unreadable, please increase its font size.
• There are some typos and phrases that need rewriting.
\end{quote}

\section{Response to Reviewer \#3}

\section{Response to Reviewer \#4}

\begin{quote}
This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
architecture for low cost and “potentially” massive distributed evolutionary algorithms. The main novelty
turns up from the fact that the architecture is developed in JavaScript.
However, JavaScript is well-known because of its lack of capability to deal with high performance
applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
believe in the possibility of “potentially” massive computation as sketched by the authors.
The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
section must be carefully checked and extended in order to fully validate the goodness of the proposed
architecture. At least, authors should show the behavior of the architecture in a real-world problem
demanding high computational performance.
Finally, English must be double-checked in order to correct minor typos such as “in the sense than[that]”
(in page 2).
\end{quote}

### 2015-11-16 special-issue-ci/Response to reviewers.tex

\begin{quote}
However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
organization and description of the platform, and the presentation and writing style.
Regarding the experimentation, I miss:
• a section discussing the threats to validity of the experiments
• an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
• A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

\begin{quote}
Regarding the organization of the paper, some important details regarding the improved framework
design named is described in the experimentation section. I would reorganize the contents to describe the
improved in detail in section 3, and focus on the experiments, its results and the discussion in section 4.
Another issue is that some details are described twice, for instance, the advantages of the use of web
workers. Fixing this issue can reduce the paper size and improve its readability.
Regarding the presentation:
• Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
• The axis labels and legends in most of the figures are unreadable, please increase its font size.
• There are some typos and phrases that need rewriting.
\end{quote}

\section{Response to Reviewer \#3}

\begin{quote}
1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
the system called NodIO consists of a REST server (coded in Node.js) and clients performing
evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
automatic restart of computations on the client's side can improve the performance of the system. The
state of the art is comprehensive and provides good introduction into the subject.

2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
Software" special issue. Thanks to both development of Internet technology and continuous increase in
computing power of electronic devices, including mobile devices, the web browser based distributed
computing can be a cheap alternative for grid and high-performance computing systems that require
high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
architecture presented inthe paper can be potentially used for solving many different time consuming
problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
solved on a desktop machine, what may not be convincing for a reader looking for an effective
optimization tool.

3. Correctness: The system works correctly - it is not very complex and the authors provide its source
code (this is a great plus!). Technical description of the system is also clear, however it lacks some
details like parameters of the machine used to determine baselines for the experiments or the format of
messages sent between the server and the clients. The only thing I noticed is that the authors used the
term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
CPUs". As far as I know, contemporary web browsers do not support multiple CPUs architecture.

4. Style: Paper is generally written in a clear way. Statements like "But fair is fair," (page 6), sound a bit
colloquially in a scientific paper and may be omitted. However, my main concern is the readability of the
graphs. General remark - font used for description of axes should be significantly larger. In Fig. 2 the
authors can change scale for axe Y, as the box for the population of 1024 individuals is very small.
What do the authors mean by "times"? Time in seconds? I am not sure if this Figure is necessary at all
as the description is understandable without it. In Fig.4 all graphs are similar and only the one in the
middle is analyzed, so maybe others are not necessary? The same is with Fig. 5. What does "x" stand
for in the description of axe X? (the same is with Fig. 8).

5. Length: As I mentioned above the authors should think of removing some of the unnecessary graphs
in order to improve readability. They admit themselves that it is hard to conclude from some of the
graphs (Fig. 5 and Fig. 8). Nevertheless, overall length of the paper is fine.

6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
distributed computing system, contrary to the description in the main text of the paper. However, it
lacks the information about the second version of the system called NodIO-W2 based on the Web
Worker architecture.

Comments to Authors:

1) I like the state of the art which includes comprehensive overview of the most important papers related
to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
with its characteristic provided on page 4. Namely:

 - Heterogeneity - it is true that almost all devices can participate in the web browser based
 distributed computing system, providing that a user did not disable Java Script in its browser.
 However if the authors plan to use HTML5 WebWorkers only modern web browsers can
 participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
 Windows Phone) are also excluded.

 - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
 implement evolutionary algorithm in that way that it wil automatically adopt to the problem type
 and the environment in which it is computed, the authors do not mention about how real
 adaptiveness of the system can be achieved. The same is with scalability. Only 5 different users
 that permanently participated in the experiments are not enough to assess the scalability of the
 system, what is a crucial parameter for distributed computing.

 - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
 code of evolutionary algorithm is sent to the client as a source code in Java Script, so some
 experienced user may quite simple decode what has to be computed. It can may come into
 possession of sensitive data or/and alter this code in order to disrupt the computations or make
 them worhless. The results send to the server may also be captured and/or altered without any
 problem since the transmission is not secure.

 In my opinion you should at least highlight above issues.

2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
maximum. In the experiments you counted how many times evolutionary algorithm was able to find
global maximum for the problem of size n=40. Initially you have found that if the population in an
evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run (page 4, if I understood it correctly).
Thus, what was the point of using distributed computing to solve such simple problem? Why the
authors did not perform their experiments for larger values of n, what could justify distributed problem
solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
it is still behind the desktop computer in more than 60% cases (when compared to the second
baseline). I suggest to perform additional computing experiments for some higher value of n in the
"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
to focus the paper on the behavior analysis of volunteers participating in a distributed computing
system, what may be also interesting an valuable for the potential readers.

3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In
Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
meaning probably that IPs from 16 to 29 did not performed any computations. What was the reason
for that?

For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC
systems - the number of participants is completely incomparable and also the type of problems
solved by the systems are of different nature and complexity.
\end{quote}

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds third reviewer by @pacastillo

The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2.

>>>>

The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2.
